TacticToe: Learning the Game of Tactics

### 2018-01-26 tactictoe_jv.tex
prediction: wip

%Mention explictly all the predictors that will be using this distance.
%tactics:
%theorems:
%terms:
%goal list: (product of tactics)
\begin{itemize}
\item term:\\
Predict the closest term in a tactic that appears in the goal.

\item :\\
Predict the theorem the closest to the current goal. Theorem dependencies are 
added during preselection.

\item goal:\\
Given a goal $g$, the classifier selects a list of previously solved goals $l$
ordered by their similarity to $g$. Since each solved goal is associated 
with a tactic in our database, the ordered list of goals induces a ordered  
list of associated tactics. And tactics are tried in this order during the 
proof search. Therefore, to order the tactics correctly, it is crucial to 
estimate the similarity between goals with mathematically relevant features.

\item list of goals:\\
Used for Monte Carlo Tree Search evaluation. Predict the closest list of goals
to the list of goal produced by a tactic.
\end{itemize}

@@

\subsection{Feature distance}\label{sec:predictions}

We estimate the similarity (or co-distance) between an open goal $g_o$ and a 
previously recorded goal $g_p$ using their respective feature sets $f_o$ and 
$f_p$.
The similarity function $sim$ computed by the k-NN algorithm is analogous 
to the
one used in the premise selection task~\cite{ckju-pxtp13}. The main idea 
is to find the features shared by the two goals and estimate the rarity of 
those features calculated via the TF-IDF~\cite{Jones72astatistical} heuristics.
In a second co-distance $sim_2$, we additionally take into account 
the
total number of 
features to reduce the seemingly unfair advantage of big feature sets in the 
first scoring function.
\[sim_1 (f_o, f_p) = {\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{\tau_1}}}\]
\[sim_2 (f_o, f_p) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{\tau_1}}}}
{ln (e + card(f_o) + card(f_p))}
\]

\begin{remark}
We use $\tau_1 = 6$ because it was found optimal in previous experiments.
\end{remark}

In addition to predicting tactics, we also used these similarity function to 
predict theorems and terms. Suprisingly, the first similarity function $sim_1$
is best to predict tactics and theorems. It is probably because there exists 
monotonocity heuristic where if a tactic can be effective on a large goal, it 
will be also effective on smaller goals that contain subterms of the larger 
goal. The same heuristic can be applied for theorems, when a large theorem 
contains more information than a theorem which is a subformula of the larger 
theorem. Nevertheless, we want to be able to predict variables (which are small 
terms) during term argument prediction. So in this case, weight down big terms, 
as implemented by the $sim_2$ function, should have a beneficial impact.
The distance $sim_2$ is also used to evaluate list of goals for proof node 
selection in the Monte Carlo tree search(MCTS) algorithm described in 
Section~\ref{sec:proofsearch}.

>>>>

%Mention explictly all the predictors that will be using this distance.
%tactics:
%theorems:
%terms:
%goal list: (product of tactics)
\paragraph{term}
Predict the closest term in a tactic that appears in the goal.

\paragraph{theorem}
Predict the theorem the closest to the current goal. Theorem dependencies are 
added during preselection.

\paragraph{tactic via goal}
Given a goal $g$, the classifier selects a list of previously solved goals $l$
ordered by their similarity to $g$. Since each solved goal is associated 
with a tactic in our database, the ordered list of goals induces a ordered  
list of associated tactics. And tactics are tried in this order during the 
proof search. Therefore, to order the tactics correctly, it is crucial to 
estimate the similarity between goals with mathematically relevant features.

Whenever we mention that we want to predict a goal, we 
actually intend to predict its associated tactics.

\paragraph{list of goals}
Used for Monte Carlo Tree Search evaluation. Predict the closest list of goals
to the list of goal produced by a tactic.

@@

\subsection{Similarity}\label{sec:predictions}

We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$ 
through their respective feature sets $f_1$ and $f_2$.
%The similarity function $sim$ computed by the k-NN algorithm is analogous 
%to the one used in the premise selection task~\cite{ckju-pxtp13}.
The estimation is based on the frequency of the features in the intersection of 
$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features 
are, the more similar two goals should be. This can be realized by 
calculating TF-IDF weights for each feature~\cite{Jones72astatistical}.
We additionally raise the weights to the sixth power giving even more credence 
to rare features which was found experimentally optimal in \cite{cezary_cite}.

The first co-distance $sim_1$ adds a modified weight of each shared 
features to compute the total score.
In a second co-distance $sim_2$, we additionally take into account 
the total number of features to reduce the seemingly unfair advantage of big 
feature sets in the first scoring function.
\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{6}}}\]
\[sim_2 (f_o, f_p) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{6}}}}
{ln (e + card(f_o) + card(f_p))}
\]

Making prediction for an object $o$ consists sorting a set of objects 
by their similarity to $o$.
\todo{mention preselection and which dataset of objects?}

We use $sim_1$ for goals and theorems and $sim_2$ for terms and list of goals.
The similarity $sim_1$ is used for predicting a tactic (via goals) because if a
tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
use of $sim_1$ in theorems. Indeed, in practice, a large theorem is often a 
conjunction of small theorems from the same domain. And if a conjunction 
contains a formula related to a problem, then the other conjuncts 
from the same domain may also help.
